---
title: "ROC curve and Feature Selection"
author: "Fakhraddin Jaf"
date: "March 18, 2017"
output: html_document
---
------------------

### Selection of features (in this case: finding the most important variables) to be used in a logistic regression model for breast cancer relapse prediction and estimating its accuracy

##### loading required libraries:
```{r chunk_name, message=FALSE, warning=FALSE}
library(caret)
library(pROC)
```
***
#### Step 1) Loading Data 
```{r}
Data_Orig <- read.table(file="datos_icb.txt", header=TRUE)
```
***
#### Step 2) A model is trained using a 10-fold cross validation
```{r message=FALSE, results="hide"}
model <-  train(recid ~ ., data=Data_Orig, 
                method= "lvq", 
                trControl = trainControl(
                            method = "cv", number = 10 , 
                            verboseIter = TRUE))
```
***
#### Step 3) Finding most important variables based on the trained model ( ... printing the result + plot):
```{r}
var_importance <- varImp(model)
print(var_importance)
ggplot(var_importance) + ggtitle("                                
                                 Variance Importance %")
```

***
#### Step 4) Building a logistic regression model using most important variables, then by using __predict()__ function, probability of relapse will be predicted
```{r}
glm_model <-  glm(recid ~ tam + grado + gang, data = Data_Orig, family = binomial("logit"))
prediction <- predict(glm_model,Data_Orig[,c("tam", "grado", "gang")], type = "r")
```
***
#### Step 5) calculating ROC and AUC
```{r}
roc1 <- roc(Data_Orig$recid, prediction, auc = TRUE)
roc1
plot(roc1, col = "blue", main="ROC Curve Plot")
auc1 = round(auc(roc1), 3)
legend(.3, .3, auc1, title = "AUC")
```

***
#### Step 6) calculating ACC
```{r}
Pred_code_01 <- ifelse(prediction >= 0.5, 1 , 0)
Orig_code_01 <- ifelse(Data_Orig$recid == "SI", 1, 0)
correctly_predicted <- sum(ifelse(Pred_code_01 == Orig_code_01, 1, 0))
ACC <- (correctly_predicted * 100)/nrow(Data_Orig)
ACC
```


